\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{float}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{hyperref}

\hypersetup{hidelinks}
\setstretch{1.1}

\title{Econ 5253 Problem Set 8}
\author{Last Name: Cheng}
\date{April 7, 2026}

\begin{document}
\maketitle

\section*{1. Simulated Data}
I used Python and set the random-number seed equal to 100.
The sample size is \(N = 100{,}000\) and the regressor matrix has \(K = 10\) columns.
The first column of \(X\) is a column of ones, and the remaining nine columns are independent standard normal draws.
I generated the disturbance vector as
\[
\varepsilon_i \sim N(0, 0.25),
\]
so the true standard deviation is \(\sigma = 0.5\).
The coefficient vector is
\[
\beta =
\begin{bmatrix}
1.5 & -1 & -0.25 & 0.75 & 3.5 & -2 & 0.5 & 1 & 1.25 & 2
\end{bmatrix}^{\prime},
\]
and the dependent variable is
\[
Y = X\beta + \varepsilon.
\]

\section*{2. Closed-Form OLS and Gradient Descent}
I first computed
\[
\hat{\beta}_{OLS} = (X'X)^{-1}X'Y.
\]
The resulting estimates are all very close to the data-generating values.
Across the ten coefficients, the largest absolute deviation from the true \(\beta\) vector is 0.002306.
Table \ref{tab:comparison} reports the estimates from every method side by side.

I then estimated the same coefficient vector by gradient descent using the learning rate required in the assignment, \(0.0000003\).
The algorithm converged in 426 iterations.
Its estimates are numerically indistinguishable from the closed-form solution: the largest absolute difference between the two vectors is only \(1.53 \times 10^{-11}\).

\section*{3. Numerical Optimization}
For the OLS objective function, I used two numerical optimizers in Python: L-BFGS-B and Nelder-Mead.
Both methods recovered the same solution as closed-form OLS up to numerical rounding.
The largest absolute difference from the closed-form estimate is \(2.74 \times 10^{-8}\) for L-BFGS-B and \(5.21 \times 10^{-9}\) for Nelder-Mead.
So the answers do not differ in any economically meaningful way.

I next estimated the normal linear-model maximum-likelihood problem with L-BFGS-B.
The estimated coefficient vector is identical to the closed-form OLS estimate to machine precision, which is exactly what we should expect under normal errors.
The estimated standard deviation is
\[
\hat{\sigma}_{MLE} = 0.499080,
\]
which is very close to the true value of 0.5.

\input{estimate_comparison.tex}

\section*{4. Direct Regression Call}
As the easy-way estimator, I ran the regression by calling the outcome vector and regressor matrix directly in Python with \texttt{statsmodels}, which is the Python analogue of using \texttt{lm(Y \textasciitilde X - 1)} without first building a data frame.
The regression output is reported in Table \ref{tab:ols}.

\input{regression_table.tex}

The direct regression call produces the same coefficient vector as the closed-form solution up to machine precision.
Relative to the ground-truth \(\beta\), every estimate is extremely close because the sample is large.
For example, the intercept estimate is 1.4991 instead of 1.5, the coefficient on \(x_5\) is 3.4992 instead of 3.5, and the coefficient on \(x_{10}\) is 1.9977 instead of 2.0.
The largest absolute gap between the direct OLS estimate and the true parameter is again only about 0.0023.

\section*{5. Reproducibility}
All computations are reproduced by \texttt{PS8\_Cheng.py}.
Running
\begin{quote}
\texttt{python PS8\_Cheng.py}
\end{quote}
from the \texttt{PS8} folder regenerates the two LaTeX tables used in this document:
\begin{quote}
\texttt{estimate\_comparison.tex}\\
\texttt{regression\_table.tex}
\end{quote}

\end{document}
