\documentclass[12pt]{article}

% ----- Packages -----
\usepackage[margin=1in]{geometry}
\usepackage{setspace}

% ----- Spacing -----
\setstretch{1.15}

\begin{document}

\title{Example Project}
\author{}
\date{}
\maketitle

My academic interests lie at the intersection of economics, accounting, and data science. I am especially interested in how capital markets react to information and economic events, such as corporate disclosures, earnings announcements, and various types of distributions. I enjoy thinking about research questions that combine economic intuition with large-scale data analysis. For me, data science is not only a technical skill but also a tool that allows me to turn ideas into testable evidence.

I want to take this class because I hope to strengthen my data and programming skills so that I can better develop and execute my own research projects. Many of my research ideas require working with large financial databases and cleaning complex datasets, so learning more efficient and rigorous data methods is very important for me. I would like to become more comfortable with coding, data management, and empirical analysis.

For the class project, I am interested in studying how the capital market reacts to certain firm-level events or distributions, such as dividends, using real financial data. I hope to design a project that connects economic theory with practical data analysis.

My goal for this class is to build strong technical skills that I can directly apply to my future research. After graduation, I plan to pursue an academic career and become an accounting professor, conducting empirical research and teaching students in accounting.


\section*{Equation}

\begin{equation}
a^2 + b^2 = c^2
\end{equation}


\end{document}
